Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-08-14 snapshot_system/abs.tex

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots and
signature-based deduplication of data blocks is necessary to eliminate excessive 
redundant blocks.  Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
% when their signatures are identical. 
however it is resource expensive and less fault-resilient to perform a global deduplication
and let a data block share by many virtual machines. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine
and associates  underlying file blocks with one VM for most of cases.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resource. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing a high deduplication 
efficiency while sustaining a good backup throughput.

>>>>

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots and
signature-based deduplication of data blocks is necessary to eliminate excessive 
redundant blocks.  Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
% when their signatures are identical. 
however it is resource expensive and less fault-resilient to perform a global deduplication
and let a data block share by many virtual machines. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine
and associates  underlying file blocks with one VM for most of cases.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resource. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing a competitive deduplication 
efficiency while sustaining a good backup throughput.

### 2013-08-14 snapshot_system/challenge.tex
revised

%In a virtualized cloud environment such as ones provided by Amazon EC2\cite{AmazonEC2} and Alibaba Aliyun\cite{Aliyun}, 
At a cloud cluster node, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks 
represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.  
As discussed earlier, collocating a backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the un-deduplicated
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architecture where
each physical runs a backup service and a distributed file system (DFS)~\cite{GFS2004,Hadoop} 
serves a backup store  for the snapshots.
The previous study shows that 
 deduplication can compress the backup copies 
effectively in a 10:1 or even 15:1 rage. 
Therefore  the portion of space in a cluster
allocated for snapshots of data should not dominate
the cluster storage usage.

@@

%The previous work for storage backup has extensively studied  data deduplication techniques that
%can eliminate redundancy globally among different files from different users.
%Offline deduplication is 
%used in ~\cite{EMC,NetAppOffline} to remove previously written duplicates during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Since it is expensive to compares a large number of chunk signatures for deduplication,
several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%As discussed in Section~\ref{sect:intro},
%there is no dedicated resource for deduplication in our targeted setting and low memory usage
% is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirements with the tradeoff of a reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
%In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.

Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,idedup}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault tolerance in the cluster-based environment that we have considered
in this paper.
%We will describe the motivation of using the cluster-based   approach
%for running the backup service and then present our solution  with fault isolation.
%In our work, the foucs 
%there is a waiting time for many duplicate detection requests. This relaxation is acceptable because 
%in our context, finishing the backup of required VM images within a reasonable time window is more
%important than optimizing individual VM block  backup requests.

@@

Our key design consideration is VM dependence minimization during deduplication and file block management.
\begin{itemize}
\item {\em Deduplication localization.}
%Previous work in deduplication~\cite{extreme_binning09,sparseindex09} has
%studied the solutions to reduce memory requirements.
%One important issue not considered in the previous work is how 
Because a data chunk is compared with signatures collected from all VMs during
the deduplication process, only one copy of duplicates is stored in the storage,
this artificially creates data dependency among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen periodically 
in a large-scale cloud and
%failure of a shared block affects many VMs.
%the loss of a shared chunks can affect many VM users whose snapshots share this 
%data block. 
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discouraged. 
Another disadvantage of sharing is that it complicates snapshot deletion, 
which  occurs frequently when snapshots expire regularly and leads a cost concern. 
The mark-sweep approach~\cite{mark-sweep} is effective for deletion, and its main cost
is to count if a data chunk is still shared by other snapshots.
Localizing deduplication can  minimize data sharing and simplify deletion while sacrificing 
deduplication efficiency, and  can faciltate parallel execition of snaptshot operations.
Thus we seek for a tradeoff.

% since storage for shared identical content chunks can become a failure point.
%Without any control of such data sharing, we can only increase  
%replication for global datasets to enhance the availability,
%but this incurs significantly higher costs.
%Thus a key goal  in our design
%is restricting the impact of  data loss to a limited number of VMs. 
%It is not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process. 
\item{\bf Management of file system blocks.}
The file block size in a distributed file system such as  Hadoop and GFS is uniform and large (e.g.  64MB),
 a data chunk in a deduplication service is of a nonuniform size with 4KB or 8KB on average.
%We need to build an intermediate layer that supports large snapshot writing with append operations
%and infrequent snapshot access.
Packaging data chunks to a file system block can create more data dependence among VMs
since a file block can be shared with even more VMs.
Thus we need to consider a minimum association of a fie system block to VMs in the packaging process.
%The distributed file system~\cite{GoogleFS,Hadoop} provides the replication support
%and we can leverage existing open-source software to implement such a layer.
%When the number of machines failed exceeds the replication degree, there could be
% some data loss.  
%Storage deduplication worsens the impact of such faults
%because  the majority of each snapshot (e.g. over 90\%) is deduplicated
%and the corresponding data chunks stored in the underlying file blocks
%are shared by many VMs. 
%The failure of one data chunk or block can impact many sharing VMs. 
\end{itemize}

@@

We call the classical deduplication approach as   VM-oblivious (VO)
because they compare fingerprints of snapshots without consideration of VM.
With the above  considerations in mind, we study a 
VM-centric approach (called VC)
for a collocated backup service with resource usage friendly
to the existing applications.  
%We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation, and then present
%an architecture and implementation design with deletion support.
%The traditional approach compares all fingerprints and
%stores one copy for all of a chunk's duplicates.
In designing a VC duplication algorithm, we have considered and adopted some of
the following previously-developed techniques.
\begin{itemize}
\item {\bf Version-based change detection}.
VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized. 
%Active snapshots will contain all the information needed to restore the virtual disk
%and when deleting a snapshot, data referenced by other snapshots are removed.
We are seeking for a tradeoff since 
cross-VM  signature comparison can deliver additional reduction~\cite{Guo2011,Dong,ExtremeBining}.
%We are seeking for additional optimization to improve duplication efficiency and design the 
%association of data chunks with the underlying file blocks so that
%file block sharing among VMs is minimized.
\item {\bf Stateless  Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing~\cite{Dong20??}
that divides the deduplication work with an approximation. This work 
is similar to Exreme Binning\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a bloomer filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, we can have adopted some of ideas to build
bins in a VM to faciliate VM-specific local search.

>>>>

%In a virtualized cloud environment such as ones provided by Amazon EC2\cite{AmazonEC2} and Alibaba Aliyun\cite{Aliyun}, 
At a cloud cluster node, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks 
represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.  
As discussed earlier, collocating a backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the un-deduplicated
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architecture where
each physical runs a backup service and a distributed file system (DFS)~\cite{GFS2004,Hadoop} 
serves a backup store  for the snapshots.
%The previous study shows that 
% deduplication can compress the backup copies 
%effectively in a 10:1 or even 15:1 rage. 
%Therefore  the portion of space in a cluster
%allocated for snapshots of data should not dominate
%the cluster storage usage.
%perhaps briefly discuss compression is high, incremental change is small from versions..???

@@

%The previous work for storage backup has extensively studied  data deduplication techniques that
%can eliminate redundancy globally among different files from different users.
%Offline deduplication is 
%used in ~\cite{EMC,NetAppOffline} to remove previously written duplicates during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Since it is expensive to compares a large number of chunk signatures for deduplication,
several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%As discussed in Section~\ref{sect:intro},
%there is no dedicated resource for deduplication in our targeted setting and low memory usage
% is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirements with the tradeoff of a reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
%In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,idedup}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault tolerance in the cluster-based environment that we have considered
in this paper.
%We will describe the motivation of using the cluster-based   approach
%for running the backup service and then present our solution  with fault isolation.
%In our work, the foucs 
%there is a waiting time for many duplicate detection requests. This relaxation is acceptable because 
%in our context, finishing the backup of required VM images within a reasonable time window is more
%important than optimizing individual VM block  backup requests.

@@

Our key design consideration is VM dependence minimization during deduplication and file block management.
\begin{itemize}
\item {\em Deduplication localization.}
%Previous work in deduplication~\cite{extreme_binning09,sparseindex09} has
%studied the solutions to reduce memory requirements.
%One important issue not considered in the previous work is how 
Because a data chunk is compared with signatures collected from all VMs during
the deduplication process, only one copy of duplicates is stored in the storage,
this artificially creates data dependency among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen periodically 
in a large-scale cloud and
%failure of a shared block affects many VMs.
%the loss of a shared chunks can affect many VM users whose snapshots share this 
%data block. 
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discouraged. 
Another disadvantage of sharing is that it complicates snapshot deletion, 
which  occurs frequently when snapshots expire regularly. 
The mark-sweep approach~\cite{mark-sweep} is effective for deletion, and its main cost
is to count if a data chunk is still shared by other snapshots.
Localizing deduplication can  minimize data sharing and simplify deletion while sacrificing 
deduplication efficiency, and  can faciltate parallel execition of snaptshot operations.
%Thus we seek for a tradeoff.

% since storage for shared identical content chunks can become a failure point.
%Without any control of such data sharing, we can only increase  
%replication for global datasets to enhance the availability,
%but this incurs significantly higher costs.
%Thus a key goal  in our design
%is restricting the impact of  data loss to a limited number of VMs. 
%It is not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process. 
\item{\em  Management of file system blocks.}
The file block size in a distributed file system such as  Hadoop and GFS is uniform and large (e.g.  64MB),
 a data chunk in a typical deduplication system is of a nonuniform size with 4KB or 8KB on average.
%We need to build an intermediate layer that supports large snapshot writing with append operations
%and infrequent snapshot access.
Packaging data chunks to a file system block can create more data dependence among VMs
since a file block can be shared with even more VMs.
Thus we need to consider a minimum association of a fie system block to VMs in the packaging process.
%The distributed file system~\cite{GoogleFS,Hadoop} provides the replication support
%and we can leverage existing open-source software to implement such a layer.
%When the number of machines failed exceeds the replication degree, there could be
% some data loss.  
%Storage deduplication worsens the impact of such faults
%because  the majority of each snapshot (e.g. over 90\%) is deduplicated
%and the corresponding data chunks stored in the underlying file blocks
%are shared by many VMs. 
%The failure of one data chunk or block can impact many sharing VMs. 
\end{itemize}

@@

We call the traditional deduplication approach as   VM-oblivious (VO)
because they compare fingerprints of snapshots without consideration of VM.
With the above  considerations in mind, we study a 
VM-centric approach (called VC)
for a collocated backup service with resource usage friendly
to the existing applications.  
%We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation, and then present
%an architecture and implementation design with deletion support.
%The traditional approach compares all fingerprints and
%stores one copy for all of a chunk's duplicates.
In designing a VC duplication algorithm, we have considered and adopted some of
the following previously-developed techniques.
\begin{itemize}
\item {\bf Version-based change detection}.
VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized. 
%Active snapshots will contain all the information needed to restore the virtual disk
%and when deleting a snapshot, data referenced by other snapshots are removed.
We are seeking for a tradeoff since 
cross-VM  signature comparison can deliver additional compression~\cite{Guo2011,Dong,ExtremeBining}.
%We are seeking for additional optimization to improve duplication efficiency and design the 
%association of data chunks with the underlying file blocks so that
%file block sharing among VMs is minimized.
\item {\bf Stateless  Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing~\cite{Dong20??}
that divides the deduplication work with an approximation. This work 
is similar to Exreme Binning\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a bloomer filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.

### 2013-08-14 snapshot_system/intro.tex

This paper seeks for a low-cost architecture option  that collocates
a backup service with other cloud services and  uses a minimum amount of resources. 
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized.
%The dirty bit approach~\cite{??}  which checks the version difference can effectively remove
%duplicates while using content fingerprints can significantly compress more~\cite{Bottlneck08}.
%On the other hand, comparing fingerprints  adds significant  memory cost. 
%Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently.
%Cloud providers often wish that the backup service only consumes  small or modest resources
%with a minimal impact to the existing cloud services.  
%We study and evaluate the integration of several deduplication strategies to meet
%the low-cost collocation requirement. 
%A recent study using subsampling~\cite{Guo2011}  can significantly reduce the memory requirement.
We also consider the fact that after
deduplication, most data chunks are shared by several to many virtual machines.
Failure of shared data chunks can have a catastrophic effect and many
snapshots of virtual machines would be affected.
The previous work in deduplication focuses on the efficiency and approximation of
finger print comparison, and has not addressed fault tolerance together with deduplication.
Thus we also seek deduplication options that yield better fault isolation.
% along with a low-cost solution.
Another issue considered is that
that deletion of old snapshots compete for computing resource as well. 
%That is because data dependence created deletion of old snapshots competes for computing resources.
% as well so deletion needs to be considered in system design. 
sharing of data chunks among oby multiple VMs and their snapshots needs to be detected during
snapshot   deletion and such dependenec complicates deletion operations. 
%The additional complexity for deletions
%come from the dependencies created by multiple links to the same problem, so
%some solution is needed to determine blocks with no dependent snapshots after deletion.
%especially when  VMs can migrate around in the cloud.

@@

The paper studies and evaluates  an integrated approach which uses  multiple duplicate detection strategies
based on  version  detection, inner VM duplicate search,
and controlled cross-VM comparison. 
This approach is VM centric by localizing duplicate detection within each VM  
and by packaging data chunks from the same VM into a file system block as much as possible.
By narrowing duplicate sharing within a small percent of data chunks,
this scheme can afford to allocate extra replicas of these shared chunks for better
fault resilience.
Localization also brings the benefits of greater ability to exploit parallelism so
backup operations can run simultaneously without a central  bottleneck.
This  VM-centric solution uses  a small amount of  memory while delivering a reasonable deduplication efficiency. 
%Our deletion strategy uses a double Bloom filter strategy for periodic mark-and-sweeping of expired data.
We have developed a prototype system that runs a cluster of Linux machines with Xen and uses 
a standard distributed file system for the backup storage. 
% with data replication and block packaging.
%and we allocate more  replicas for the commonly-shared  data 
%blocks among VM snapshots to enhance fault tolerance.

>>>>

This paper seeks for a low-cost architecture option  that collocates
a backup service with other cloud services and  uses a minimum amount of resources. 
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized.
%The dirty bit approach~\cite{??}  which checks the version difference can effectively remove
%duplicates while using content fingerprints can significantly compress more~\cite{Bottlneck08}.
%On the other hand, comparing fingerprints  adds significant  memory cost. 
%Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently.
%Cloud providers often wish that the backup service only consumes  small or modest resources
%with a minimal impact to the existing cloud services.  
%We study and evaluate the integration of several deduplication strategies to meet
%the low-cost collocation requirement. 
%A recent study using subsampling~\cite{Guo2011}  can significantly reduce the memory requirement.
We also consider the fact that after
deduplication, most data chunks are shared by several to many virtual machines.
Failure of few shared data chunks can have a 
%catastrophic 
broad effect and many
snapshots of virtual machines would be affected.
The previous work in deduplication focuses on the efficiency and approximation of
finger print comparison, and has not addressed fault handling issues  together with deduplication.
Thus we also seek deduplication options that yield better fault isolation.
% along with a low-cost solution.
Another issue considered is that
that deletion of old snapshots compete for computing resource as well. 
%That is because data dependence created deletion of old snapshots competes for computing resources.
% as well so deletion needs to be considered in system design. 
Sharing of data chunks among by multiple VMs needs to be detected during
snapshot   deletion and such dependence complicates deletion operations. 
%The additional complexity for deletions
%come from the dependencies created by multiple links to the same problem, so
%some solution is needed to determine blocks with no dependent snapshots after deletion.
%especially when  VMs can migrate around in the cloud.

@@

The paper studies and evaluates  an integrated approach which uses  multiple duplicate detection strategies
based on  version  detection, inner VM duplicate search,
and controlled cross-VM comparison. 
This approach is VM centric by localizing duplicate detection within each VM  
and by packaging data chunks from the same VM into a file system block as much as possible.
By narrowing duplicate sharing within a small percent of common data chunks and exploiting their popularity,
this scheme can afford to allocate extra replicas of these shared chunks for better
fault resilience while sustaining competitive deduplication efficiency.
Localization also brings the benefits of greater ability to exploit parallelism so
backup operations can run simultaneously without a central  bottleneck.
This  VM-centric solution uses  a small amount of  memory while delivering a reasonable deduplication efficiency. 
%Our deletion strategy uses a double Bloom filter strategy for periodic mark-and-sweeping of expired data.
We have developed a prototype system that runs a cluster of Linux machines with Xen and uses 
a standard distributed file system for the backup storage. 
% with data replication and block packaging.
%and we allocate more  replicas for the commonly-shared  data 
%blocks among VM snapshots to enhance fault tolerance.
